---
title: "Energy expenditure share"
author: "Shiya Zhao"
date: "First made: 2024/12/19 </br> Last update: `r format(Sys.Date(), '%Y/%m/%d')`"
output: 
  html_document:
    theme: lumen
    number_sections: false
    highlight: tango
    df_print: paged
    toc: true
    toc_depth: 3
    toc_float: true
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Desktop/Model/AIMPHI/YSSP_AIMPHI/exe")

require(tidyverse) 
require(doParallel)
require(foreach)
require(ggpmisc)  %>%
  suppressWarnings()
require(ggplot2)  %>%
  suppressWarnings()
require(knitr)
require(ggthemes)
require(fitdistrplus)
require(pscl)
detach("package:pscl", unload = TRUE)
getwd()

```


```{r}

source(paste0(dir_ModelComparison, "/inc_prog/0_Directories.R"))
source(paste0(dir_ModelComparison, "/inc_prog/0_PlotSettings.R"))
source(paste0(dir_ModelComparison, "/inc_prog/igdx_GAMS_PATH.R"))
source(paste0(dir_ModelComparison, "/inc_prog/0_Maps_global.R"))
```
```{r 1-population, echo=FALSE}
df_pop <- paste0(dir_data,"/Inputdata_base.gdx")%>% rgdx.param("Population") %>% 
  rename.vars(from = colnames(.), to = c("Ref", "R",  "Y", "Pop")) %>% 
  filter(Y %in% c(seq(2020, 2090, 10)), Ref == "SSP2") %>% dplyr::select(-Ref) 

```

```{r 1-country-list, echo=FALSE}
  df_FPL_wd <- readxl::read_excel(paste0(dir_ModelComparison, "/data/WorldBank/OGHIST.xlsx"), sheet = "Country Analytical History")
  colnames(df_FPL_wd) <- df_FPL_wd[4,]
  colnames(df_FPL_wd)[1] <- "R"
  df_FPL_wd <- df_FPL_wd %>% dplyr::select(R, FY21)
  df_FPL_wd <- df_FPL_wd[-1:-10,]
  
  df_FPL_wd <- df_FPL_wd %>% mutate(adjust = case_when(FY21 == "L" ~ .63, FY21 == "LM" ~ .56, FY21 == "UM" ~ .46, FY21 == "H" ~ .46)) %>% dplyr::rename(category = FY21) %>% filter(!is.na(R))
  head(df_FPL_wd)
```

## objectives
This document is to show the energy expenditure share of different households in different countries using decile level projections. By deriving the relationship between household income and energy expenditure, we can estimate energy poverty indicators in different countries and time using the energy expenditure share of different households in different countries.  

Three energy poverty indicators are considered in this document:
1. Low-income-high-energy-expenditure (LIHEE) rate
2. Hidden energy poverty rate
3. 2M indicator
But first, we need to derive the relationship between household income and energy expenditure.

```{r data}

load(paste0(dir_output_YSSPpaper, "/Rdata/5_df_ExpShare.RData"))
df_ExpShare_ene <- df_ExpShare %>% filter(I_abb == "Energy") %>% 
  left_join(df_budget %>% dplyr::rename(budget = value)) %>% distinct()

pdata <- df_ExpShare_ene %>% 
  filter(Y %in% c(seq(2020, 2090, 10)), target == "1.5C", R== "ZAF") 

p <- ggplot(pdata, aes(x = budget, y = value * 100, color = paste0(model, revenue), shape = gini)) + 
  geom_point() + 
  # geom_smooth(method = "lm", se = FALSE) + 
  theme_minimal() + 
  theme(legend.position = "right") + 
  labs(x = "Household income", y = "Energy expenditure share") + 
  facet_wrap(target~Y, scales = "free_y") + 
  scale_color_viridis_d(option = "D", begin = 0.1, end = 0.9)
p

```

## Fit models 

```{r fit-lognorm, echo=FALSE}
colnames(df_ExpShare_ene)
df0 <- df_ExpShare_ene %>% 
  filter(Y %in% c("2080", "2070", "2050", "2040", "2030", "2020")) %>% # R %in% c("CHN", "JPN", "ZAF", "IND") c(seq(2070, 2080, 10))
  mutate(group = paste0(model, "_", R, "_", Y, "_", target, "_", revenue), 
         # group = paste0(model, "_", R, "_", Y, "_", target, "_", revenue), 
         budget = budget/1000) 
# budget thousand dollors

# switch function ----------------------------------------------
sw_fun <- "log-inverse"
# sw_fun <- "inverse"
# sw_fun <- "exp" # fail to converge in some cases



lis_group <- unique(df0$group)
df1 <- data.frame()
# parallel calculation
  time3 <- Sys.time()
  cluster <- makeCluster(10)
  registerDoParallel(cluster)
  df1 <- foreach(i = 1:length(lis_group),.combine=rbind, .packages = c('tidyverse', 'gdxrrw')) %dopar% {
    MacOS <- gdata::startsWith(Sys.getenv("R_HOME"), "/")
    if(MacOS){
      igdx("/Library/Frameworks/GAMS.framework/Versions/44/Resources")
      print("MacOS")
    }else{
      print("not MacOS")
    }

  df <- df0 %>% 
    filter( group %in% lis_group[i])
    if(sw_fun == "log-inverse"){
      
      m_reg <- nls(value~ a/(1+log(budget+1)) + b,
               data=df,
               start = list(a = .1, b = .05),   
               lower = c(a = 0.001, b = .001),
               upper = c(a = .8, b = .8),
               algorithm = "port")  # initial values

      df$fitted <- predict(m_reg)
      df$a <- coef(m_reg)["a"] # slope
      df$b <- coef(m_reg)["b"] # intercept
      rss <- sum((df$value - df$fitted)^2) # Residual sum of squares
      tss <- sum((df$value - mean(df$value))^2) # Total sum of squares
      df$Rsqr <- 1 - rss/tss # R squared
           
    }else if(sw_fun == "inverse"){
            m_reg <- nls(value~ a/(budget+1) + b,
               data=df,
               start = list(a = .1, b = .05),   
               # lower = c(a = 0, b = .01),
               # upper = c(a = .3, b = .3), 
               algorithm = "port")  # initial values
               
      df$fitted <- predict(m_reg)
      df$a <- coef(m_reg)["a"] # slope
      df$b <- coef(m_reg)["b"] # intercept
      rss <- sum((df$value - df$fitted)^2) # Residual sum of squares
      tss <- sum((df$value - mean(df$value))^2) # Total sum of squares
      df$Rsqr <- 1-rss/tss # R squared
    
    }else if(sw_fun == "exp"){
      # if we use the exponential model
      m_reg <- nls(value ~ a* exp(-k*budget ) + c,
                 data = df,
                 start = list(a = .1, k = 1/1000,  c = .05),   # initial values
                 lower = c(a = 0, k = 0,  c = .01),
                 upper = c(a = .3, k = +Inf, c = .1), 
               algorithm = "port")  # Lower bounds for a, b, c
      df$fitted <- predict(m_reg)
      df$a <- coef(m_reg)["a"]
      df$k <- coef(m_reg)["k"]
      df$c <- coef(m_reg)["c"]
      rss <- sum((df$value - df$fitted)^2) # Residual sum of squares
      tss <- sum((df$value - mean(df$value))^2) # Total sum of squares
      df$Rsqr <- 1-rss/tss # R squared
    }
  return(df)
  
  }
  stopImplicitCluster()
  stopCluster(cluster)
  time4 <- Sys.time()
  print(time4-time3)
  
df_para <- df1 %>% dplyr::select(model, R, Y, target, revenue, a, b) %>% mutate(type = sw_fun) %>% distinct()

save(df_para, df0, df1, file = paste0(dir_output_YSSPpaper, "/RData/5_7_EneExpShare_para.RData"))


pdata <- df1 %>% dplyr::select(-"scenario") %>% left_join(df_FPL_wd) %>% filter(!is.na(category))
p <- ggplot() +
  geom_point(data = pdata, aes(x = fitted, y = value, color = R_CGE, shape = gini)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed", color = "grey") +
  ggthemes::theme_few() + facet_wrap(~category, scales = "free") 
p


pdata <- df1 %>% pivot_longer(cols = c(value, fitted), names_to = "type", values_to = "value") %>% dplyr::select(-"scenario") %>% left_join(df_FPL_wd) %>% filter(!is.na(category))
p <- ggplot() +
  geom_point(data = pdata, aes(x = budget, y = value, color = type, shape = gini)) +
  ggthemes::theme_few() + facet_wrap(~category, scales = "free") 
p

p <- ggplot() +
  geom_boxplot(data = pdata, aes(x = category, y = Rsqr), outliers = FALSE) +
  ggthemes::theme_few() + facet_wrap(~target)
p


```
# Calculating the LIHC poverty rate
Having the income thresholds for energy poverty related indicators, we can calculate the rate and population of energy poverty in different countries and years.  
```{r}
# load the income distribution parameters
for(m in seq_along(lis_model)){
  # m <- 1
  if(lis_model[m] == "AIM-Hub"){
      gdx_in <- AnaExp
    }else if(lis_model[m] == "MESSAGEix"){
      gdx_in <- AnaExpmsg
    }  
  
  df_result0 <- df_result %>% filter(model == lis_model[m])
  
  df_mu <- rgdx.param(gdx_in, "Mu_exp")
  
  df_sigma <- rgdx.param(gdx_in, "Sigma_exp")

  df_result1 <- df_mu %>% left_join(df_sigma) %>% F_reve_mutate() %>% full_join(df_result0 %>% dplyr::select(-a, -b)) %>% mutate(rate)
}



```

# Calculate the energy poverty indexes
## Income threshold for Low-income-high-consumption indicator based on the concrete model
```{r inc-LIHC}
# define the function

F_eq_LIHC <- function(x, a, b){
  return(x*(1-b-a/( 1+log(x+1) ) ))
}
# val_target <- 2.15*365

result <- sapply(1:nrow(df_para), function(i){
  # extract parameters
  a <- df_para[i, "a"]
  b <- df_para[i, "b"]
  
  # define the target value to solve f(x) = val_target (thousand dollors), here it is the annual income at thousand dollors
  val_target <- 0.785
  # solve the equation
  result <- uniroot(function(x) F_eq_LIHC(x, a, b) - val_target, c(0, 100))
  return(result$root)
})

df_result <- df_para
df_result$est_x <- result
df_result <- df_result %>% mutate(est_y = b+a/( 1+log(est_x+1) ))
```

# LIHC index
## Calculating the income threshold for Low-income-high-consumption indicator based on discrete segment level
```{r discrete-LIHC}
# loading the discrete segment level data

lis_y <- c(seq(2020, 2070, 10))
# lis_ref_loop <- lis_Ref
lis_R_loop <- unique(df_ExpShare_ene$R)
df_enepov_out <- data.frame()
df_enepov_out1 <- data.frame()

# apply the function
for(m in seq_along(lis_model)){
  # m <- 1
  if(lis_model[m] == "AIM-Hub"){
      gdx_in <- demandcge
    }else if(lis_model[m] == "MESSAGEix"){
      gdx_in <- demandmsg
    }  
  
  Seg_pop0 <- rgdx.param(gdx_in, "FreqSegall") %>%
    rename.vars(from = colnames(.), to = c("R", 'Ref','Y','Seg','pop')) %>%
    filter(Y %in% lis_y) %>%
    group_by(R,Ref,Y) %>%
    dplyr::reframe(pop = pop, Seg = Seg) %>% 
    SegNum()

  ExpNational_tmp0 <- rgdx.param(gdx_in, "ExpNational") %>%
    rename.vars(from = colnames(.), to = c("R", 'Ref','Seg','Y',"I",'ExpNational')) %>%
    filter( #Ref %in% lis_ref_loop,
            Y %in% lis_y,
    ) %>%
    dplyr::group_by(R, Ref, Seg, Y) %>%
    dplyr::mutate(share = ExpNational/sum(ExpNational)) %>%
    ungroup() %>%
    filter(I %in% lis_Iene)
  
  # run the loop for each country in parallel
  time3 <- Sys.time()
  cluster <- makeCluster(10)
  registerDoParallel(cluster)
  df_enepov_out1 <- foreach(r = 1:length(lis_R_loop),.combine=rbind, .packages = c('tidyverse')) %dopar% {
    # r <- 1
    Seg_pop <- Seg_pop0 %>% filter(R %in% lis_R_loop[r])
    ExpNational_tmp <- ExpNational_tmp0 %>% filter(R %in% lis_R_loop[r])
    
    F_EnePov(Seg_pop, ExpNational_tmp, PVL = 2.15,lis_ref = lis_Ref,lis_y = c(2020, 2030, 2050, 2070),lis_Iene = c("Energy"), median = F)
  
  } # END of the loop for each country
  stopImplicitCluster()
  stopCluster(cluster)
  time4 <- Sys.time()
  print(time4-time3)
  
  df_enepov_out <- df_enepov_out %>% rbind(df_enepov_out1 %>% mutate(model = lis_model[m]))

} # END of the loop for each model

# see the regions and models
unique(df_enepov_out$R) %>% length()
unique(df_enepov_out$Ref)
unique(df_enepov_out$model)

df_enepov <- df_enepov_out %>% mutate_at(vars(`TwoM`:`LIHC`), ~replace(., is.na(.), 0)) %>% left_join(df_pop) %>% mutate(HEP_pop = HEP*Pop, LIHC_pop = LIHC*Pop) %>% filter(R != "WLD") %>% F_reve_mutate() %>% left_join(df_FPL_wd) %>% filter(!is.na(category)) 

```

```{r plot}

# plot the results
pdata <- df_enepov %>% dplyr::group_by(target, gini, revenue, model,  Y) %>% reframe(HEP_pop = sum(HEP_pop), LIHC_pop = sum(LIHC_pop)) %>% filter(gini == "consistent", target %in% c("No-Miti", "1.5C")) 

p <- ggplot(pdata %>% filter(revenue == "Neutral")  %>% pivot_longer(cols = c("HEP_pop","LIHC_pop"), names_to = "Type", values_to = "value")) + 
  geom_col(aes(x = Y, y = value/1000000, fill = target),
           position = "dodge") + 
  # geom_line() + 
  ggthemes::theme_pander() + 
  theme(legend.position = "right") + 
  labs(x = "Year", y = "Low-income-high-consumption (million)") + 
  facet_wrap(Type~model)
p
```

```{r plot-LIHC}
df_LIHC <- df_enepov %>% left_join(Map_r_PHI2Hub) %>% dplyr::group_by(R_CGE, target, gini, revenue, model,  Y) %>% reframe(value = sum(LIHC_pop)) %>% F_R_CGEfull() %>% dplyr::rename(R = R_CGE) 

# regional LIHC
pdata <- df_LIHC %>% filter(target %in% c("No-Miti", "1.5C", "2C")) %>% F_change() %>% filter(gini == "consistent", Y %in% c("2030", "2040", "2050", "2070"), target == "1.5C") %>% mutate(revenue = factor(revenue, levels = c("Neutral", "EPC")))

p <- ggplot() + 
  geom_col(pdata, mapping = aes(x = Y, y = change/1000000, fill = R),
           position = "stack", width = .7) + 
  geom_abline(slope = 0, intercept = 0, color = "grey") +
  MyTheme + 
  theme(legend.position = "right") + 
  labs(x = "Year", y = "LIHC (million)", fill = "Model") + 
  scale_fill_manual(values = palette_R_CGE) +
  facet_grid(model~revenue)
p
ggsave(p, file = paste0(dir_output_YSSPpaper, "/fig/5_7_LIHC_region.pdf"), width = 15, height = 12, units= "cm")


# Global LIHC
pdata <- df_LIHC %>% filter(target %in% c("No-Miti", "1.5C", "2C")) %>% dplyr::group_by(target, gini, revenue, model,  Y) %>% reframe(value = sum(value)) %>% mutate(R = "WLD") %>% F_change() %>% filter(gini == "consistent", Y %in% c("2030", "2040", "2050", "2070"), target == "1.5C") %>% mutate(revenue = factor(revenue, levels = c("Neutral", "EPC")))

p <- ggplot() + 
  geom_col(pdata, mapping = aes(x = Y, y = change/1000000, fill = model),
           position = "dodge", width = .7) + 
  geom_abline(slope = 0, intercept = 0, color = "grey") +
  MyTheme + 
  theme(legend.position = "right") + 
  labs(x = "Year", y = "LIHC (million)", fill = "Model") + 
  scale_fill_manual(values = palette_model) +
  facet_grid(~revenue)
p
ggsave(p, file = paste0(dir_output_YSSPpaper, "/fig/5_7_LIHC_global.pdf"), width = 13, height = 7, units= "cm")
```


# Additional optimization

The analysis below is to fit the distribution to the data using more complicated/customized functions.   
It is commented out because it is not necessary for the current analysis.
```{r other-roptimization}
# # functional form ----------------------------------------------
# # Fit lognormal distribution to the 'value' column
# F_lnorm <- function(x, para = c(a1, a2)){
#   sigma <- para[1]
#   mu <- para[2]
#   
#   # project y
#   y <- 1/(x*sigma*(2*pi)^0.5) * exp(-0.5*((log(x)-mu)/sigma)^2)
#   return(y)
# }
# 
# # beta distribution
# F_beta <- function(x, para = c(a1, a2)){
#   a1 <- para[1]
#   a2 <- para[2]
#   
#   # project y
#   y <- (x^(a1-1) * (1-x)^(a2-1))/beta(a1, a2)
#   return(y)
# }
# 
# # gamma distribution
# F_gamma <- function(x, para = c(a1, a2)){
#   a1 <- para[1]
#   a2 <- para[2]
#   
#   # project y
#   y <- (x^(a1-1) * exp(-x/a2))/(gamma(a1)*a2^a1)
#   return(y)
# }
# 
# 
# # log likelihood function ----------------------------------------------
# F_logll_lnorm <- function(para){
#   df1 <- df_in %>% 
#     # distinct() %>% # "Solid", "Liquid", "Gas", "Electricity", "Bio_liq", "Bio_pri"
#     mutate(est = F_lnorm(x, para),
#            xlogP = log(est)*x)
#   
#   LL <- sum(df1$xlogP, na.rm = TRUE)
#   
#   return(LL)
# }
# 
# head(df0)
# 
# df_in <- df0 %>% dplyr::rename(x = budget, y = value) %>% dplyr::select(-I_abb, -revenue, -target) %>% filter(R == "CHN", model == "AIM-Hub", scenario == "No-Miti_Neutral")
# 
# para <- numeric(2)
# para <- c(5, 6)
# para <- c(1, 10.8)
#   df1 <- df_in %>% 
#     mutate(est = F_lnorm(x, para),
#            xlogP = log(est)*x)
# 
#   res <- optim(para, F_logll_lnorm, # lower = c(rep(-Inf, 6), rep(0,3)), upper = c(rep(0, 6), rep(Inf,3)), 
#              method = "Nelder-Mead",
#              hessian = FALSE, control = list(maxit = 600, fnscale = -1)) 
# # By default optim performs minimization, but it will maximize if control$fnscale is negative.
# 
# print(res)
# 
# 

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
